\Huge{Vipassana for Hackers} \\ \large{Paper Two: The Brain} \\ \large\textit{Version 0.2}

### 2019-09-29 paper-two/paper-two.tex
reformatted paper two with physics article template

\documentclass[twocolumn]{article}

\usepackage{polyglossia}
\setdefaultlanguage{english}
\setotherlanguage{sanskrit}
\usepackage{fontspec}
\newfontfamily\devanagarifont[Scale=MatchUppercase]{Kohinoor Devanagari}

\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{float}
\usepackage[bottom]{footmisc}
\usepackage{enumitem}
\usepackage{hyperref}

\setlength{\parindent}{1.3em}
\setlength{\parskip}{0.7em}
\setlist{noitemsep}

@@

\begin{titlepage}
   \vspace*{\stretch{1.0}}
   \begin{center}
     \Huge\textbf{Vipassana for Hackers}\\
     \Huge{Paper Two: The Brain}\\
     \vspace{5cm}
     \large\textit{Preethi Govindarajan}\\
     \large\textit{Version 0.1}\\
     \large\textit\today\\
     \vspace{5cm}
     \large\textit{Contributions by: Steven Deobald}\\
   \end{center}
   \vspace*{\stretch{2.0}}
\end{titlepage}

@@

\begin{quote}
  \textbf{The brain itself is just a physical organ. As you deal with other parts of
    the body, you deal with the brain in the same way, that's all. Nothing special to
    do with the brain. But the mind is totally different. In the West, all importance
    is given to the brain as if the mind is located here. Nothing doing, it is
    everywhere. The mind is in the whole body. So give attention to the whole
    body.} \cite{goenkabrain}
\end{quote}

@@

[ TODO: insert switching diagram ]

@@

[ TODO: MRI of DMN image ] \cite{dmnmri}

@@

[ TODO: explain IPL? ]

It is hypothesized that the activation of these three regions of the brain (PCC,
vmPFC, and IPL) are responsible for providing a sense of self --- a subjective agent
in space and time. It functions as a network in which phenomena such as self
representations are accessible via the PCC [ TODO: ref ] and are gated into conscious awareness by
activity in the mPFC, as influenced by changing internal and external demands. The
PCC acts as a brain-wide connectivity hub, through which a unitary notion of self is
created by a large scale integration of DMN activity. \cite{mappingself}

@@

[ TODO: expand? how? ]

[ TODO: Summarize DMN? Maybe introduce why it's relevant to meditation? ]

@@

[ TODO: image of Approximate location of the OFC on an MRI ]

@@

[ TODO: why? how? expand. ]

[ TODO: Summarize CEN? Relevance to meditation? ]

@@

The AIC and ACC together give rise to our interoceptive and conscious
self-awareness. [ TODO: reference ]

@@

The limbic system is a complex set of structures that lies on both sides of the
Thalamus, just under the Cerebrum. It includes the Hypothalamus, the Hippocampus, the
Amygdala, and several other nearby areas. It appears to be primarily responsible for
our emotional life, and has a lot to do with the formation of memories. [ TODO:
  reference ]

@@

(The somatosensory cortex is located in the gyrus and the parietal lobes: The sensory
homunculus which is a part of the somatosensory cortex is a cortical representation
of the body based on the degree of sensory innervation) [ TODO: Maybe move this to
  the anatomy section? ]

[ TODO: sensory homunculus diagram ]
Somatosensory and Motor cortices

@@

There have been studies looking at both the structural changes in the brain and
functional changes in brain activity, especially to the DMN and the SN. But such
studies are usually only done on meditators who have clocked thousands of hours of
practice, which diverges from the purpose of this paper. [ TODO: ref? ]

@@

\pagebreak

\begin{thebibliography}{9}
\raggedright

@@

\bibitem{dmnmri}
  \textit{Magnetic Resonance Image of the areas of the brain in the Default Mode Network.}
  http://www.frontiersin.org/Neurotrauma/10.3389/fneur.2013.00016/full

@@

\bibitem{parietallobes}
  Martin, R. E.
  \textit{Let’s Get to Know the Parietal Lobes!}
  \url{http://gablab.mit.edu/downloads/Parietal_Primer.pdf} [ TODO: 404? ]

@@

>>>>

\documentclass[a4paper, amsfonts, amssymb, amsmath, reprint, showkeys, nofootinbib, twoside]{revtex4-1}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[colorinlistoftodos, color=green!40, prependcaption]{todonotes}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{float}
\usepackage[bottom]{footmisc}
\usepackage{enumitem}
\usepackage{hyperref}

\bibliographystyle{apsrev4-1}
\setlist{noitemsep}

@@

\title{%
  \Huge{Vipassana for Hackers} \\
  \large{Paper Two: The Brain} \\
  \large\textit{Version 0.2}
}
\author{Preethi Govindarajan}
\email[Correspondence email address: ]{preethi@deobald.ca}
\affiliation{Siggu.org}
\date{\today}

\begin{abstract}
  \todo{add an abstract} This is the abstract. etc. etc.
\end{abstract}

\keywords{neuroscience, vipassana, meditation}

\maketitle

\listoftodos

@@

\begin{quote}
  \textbf{The brain itself is just a physical organ. As you deal with other parts of
    the body, you deal with the brain in the same way, that's all. Nothing special to
    do with the brain. But the mind is totally different. In the West, all importance
    is given to the brain as if the mind is located here. Nothing doing, it is
    everywhere. The mind is in the whole body. So give attention to the whole
    body.} --- S.N. Goenka \cite{goenkabrain}
\end{quote}

@@

\todo{insert switching diagram}

@@

\todo{ insert image of ``MRI of DMN'' }
\textit{Magnetic Resonance Image of the areas of the brain in the Default Mode Network.}
http://www.frontiersin.org/Neurotrauma/10.3389/fneur.2013.00016/full

@@

\todo{explain IPL?}

It is hypothesized that the activation of these three regions of the brain (PCC,
vmPFC, and IPL) are responsible for providing a sense of self --- a subjective agent
in space and time. It functions as a network in which phenomena such as self
representations are accessible via the PCC \todo{add PCC reference} and are gated into conscious awareness by
activity in the mPFC, as influenced by changing internal and external demands. The
PCC acts as a brain-wide connectivity hub, through which a unitary notion of self is
created by a large scale integration of DMN activity. \cite{mappingself}

@@

\todo{expand on DMN involvement in past/future/stories. how does this work?}

\todo{Summarize DMN? Maybe introduce why it's relevant to meditation?}

@@

\todo{add image of Approximate location of the OFC on an MRI}

@@

\todo{why and how does the PCC receive external information and activate the motor cortex/DL-PFC? expand.}

\todo{Summarize CEN? Relevance to meditation?}

@@

The AIC and ACC together give rise to our interoceptive and conscious
self-awareness. \todo{add reference justifying statement that AIC+ACC create
  conscious self-awareness}

@@

The limbic system is a complex set of structures that lies on both sides of the
Thalamus, just under the Cerebrum. It includes the Hypothalamus, the Hippocampus, the
Amygdala, and several other nearby areas. It appears to be primarily responsible for
our emotional life, and has a lot to do with the formation of memories. \todo{add
  reference for limbic system}

@@

(The somatosensory cortex is located in the gyrus and the parietal lobes: The sensory
homunculus which is a part of the somatosensory cortex is a cortical representation
of the body based on the degree of sensory innervation) \todo{Maybe move sensory
  homunculus diagram and explanation to the anatomy section?}

\todo{add sensory homunculus diagram}
Somatosensory and Motor cortices

@@

There have been studies looking at both the structural changes in the brain and
functional changes in brain activity, especially to the DMN and the SN. But such
studies are usually only done on meditators who have clocked thousands of hours of
practice, which diverges from the purpose of this paper. \todo{add references with
  evidence of structural changes in the brain after meditation}

@@

\section{Acknowledgements}

Thank you to Steven Deobald for reviews, edits, and corrections.

\begin{thebibliography}{99}

\section*{References}

@@

@@

\bibitem{parietallobes}
  Martin, R. E.
  \textit{Let’s Get to Know the Parietal Lobes!}
  \url{http://gablab.mit.edu/downloads/Parietal_Primer.pdf}
  \todo{parietal lobes article returns a 404. fix.}

@@

% \appendix*
% \input{sections/appendix1.tex}
